---
title: "hookCompetition"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hookCompetition}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.width=12, fig.height=8) 
```

```{r setup}
library(hookCompetition)
library(gfiphc)
library(sf)
library(inlabru)
library(ggpolypath)
library(tidyverse)
```

```{r loadhiddenfunctions, include=F, echo=F}
colsc <- function(...) {
  ggplot2::scale_fill_gradientn(colours = rev(RColorBrewer::brewer.pal(11,"RdYlBu")),
                       limits = range(..., na.rm=TRUE))
}
```


## How to use the hookCompetition package

This markdown document will outline the steps needed to use the hookCompetition package.

# Step 1 - Setting parameters for generating the Rmarkdown document

```{r loadpackage}
# Set the number of cores to use to generate the vignette
n_cores <- 1
# Set the number of samples to use for bootstrapping for Monte Carlo approximations to 
# posterior quantities. Should be set much higher for real applications!
n_samp <- 100
```

# Step 2 - Loading the data.

The function `read_Data_hookcomp()` reads in longline catch count data collected by the IPHC. The arguments to the function are:

- `species_vec` is a character vector of the one or more species desired for modelling (e.g. 'yelloweye rockfish')
- `at_PBS` is a logical flag determining if the user has access to the gf data. If FALSE, yelloweye rockfish must be the specified species. 
- `simple_species_vec` is a character vector containing the desired shortened names of the species (e.g. 'yelloweye')
- `data_wd` is a character vector stating the working directory where the rds files will be saved
- `min_dist` is a numeric stating the minimum distance (km) required for a temporary IPHC station to be from a permanent IPHC station to be included. If 0, discard all temporary stations from analysis.
- `min_dist_to_boundary` is a numeric stating the minimum distance from a survey boundary (km) required for an IPHC station to be included as data. If 0, then only include stations strictly within `survey_boundaries`
- `years_all_vec` is a numeric vector stating the years all hooks were enumerated
- `years_20_vec` is a numeric vector stating the years only 20 hooks were enumerated. Only data from years specified in the above two arguments are downloaded and included for later analysis.
- `survey_boundaries`   survey_boundaries is a `sf` polygons object containing user-specified survey boundaries for indices to be computed in. If NULL, then QCS, WCHG, WCVI, HS is used. The data.frame *must* contain a factor variable `Region` with levels equal to the names of the regions
- `preserve_inter_regional_differences` Logical. If `TRUE` the relative magnitudes in catch rates between the subregions are preserved. If `FALSE` the catch rates are normalized separately within each subregion. Use the latter if differences in absolute abundances between regions are not desired and only temporal trends *within* each subregion are desired. 


```{r loaddata, message=F, warning=F, cache=F}
data <- hookCompetition::read_Data_hookcomp(
  species_vec = c("yelloweye rockfish",
            "arrowtooth flounder",
            "lingcod",
            "north pacific spiny dogfish",
            "pacific cod",
            "pacific halibut",
            "redbanded rockfish",
            "sablefish",
            "walleye pollock"),
  simple_species_vec = c("yelloweye",
            "arrowtooth",
            "lingcod",
            "dogfish",
            "cod",
            "halibut",
            "redbanded",
            "sablefish",
            "walleye"),
  at_PBS = F,
  data_wd = './',
  min_dist = 50,
  min_dist_to_boundary = 0,
  years_all_vec = c(1995,1996,2003:2012, 2014:2019),
  years_20_vec = c(1997:2002, 2013, 2020, 2021),
  survey_boundaries=NULL
)

```

The function `read_Data_hookcomp()` returns a list containing two entries:
- `Reduced_data_sp` is a `sf` points containing all the set-level catch counts of the species specified (variable name `N_it_yelloweye`).
- `survey_boundaries` are the survey boundaries projected into the same CRS as the `Reduced_data_sp` object.

To put the objects in the list into the current working environment, run the following:

```{r environment, message=FALSE, warning=FALSE}
list2env(data, globalenv())
rm(data)
```

*Note:* to use the package on custom data, ensure the following variable names are used inside the `sf` points object containing the catch counts:

- `N_it_species` (where species is the name of the species used (e.g. yelloweye)). This contains the catch counts of the species
- `effSkateIPHC` contains a measure of effort (e.g. number of hooks used etc.,)
- `prop_removed` contains the proportion of baits removed from the fishing event
- `year` the year of the fishing event
- `region_INLA` the region where the fishing event took place. An index (numeric variable) matching the corresponding row in `survey_boundaries`
- `station` containing the index of the fishing station where the fishing event took place. Useful if the same locations are visited each year (for multiple years)
- `twentyhooks` a bindary variable indicating years where a different number of hooks were used. Useful if 2 fishing protocols are used and differences want to be controlled for. Set to 0 if not relevant.

# Step 3 - Fit the indices without the spatio-temporal random effects. 

We derive bootstrap indices for a set of test species using the function `bootstrap_index_fun_run`. We derive an overall index and a subregion index (with subregions defined by the survey boundaries).

The function `bootstrap_index_fun_run` takes as arguments:

- `data` a `sf` points object containing the IPHC data (typically the `Reduced_data_sp` object from `read_Data_hookcomp`)
- `species` a character of the species (typically matching `simple_species_vec` from `read_Data_hookcomp()` linked to the variables `N_it` in the dataframe)
- `survey_boundaries` a `sf` polygons object containing the survey boundary definitions (typically outputted from `read_Data_hookcomp()`)
- `R` specifies the number of bootstrap samples desired
- `ncpus` specifies the number of cores to use in parallel
- `type` specifies the type of bootstrap intervals to use (default percentile `perc`)
- `return` when TRUE it returns the indices
- `plot` when TRUE it returns the plot (can be memory costly when using large numbers of species). 

```{r bootstrap, cache=F, warning=F, message=F}
bootstrap_ind <-
bootstrap_index_fun_run(species = 'yelloweye',
                          data = Reduced_data_sp,
                          survey_boundaries = survey_boundaries,
                          R=n_samp,
                          return=T,
                          ncpus=n_cores,
                          type='perc',
                          plot=T
                          )

```

The function `bootstrap_index_fun_run` returns a list containing 3 objects:

- `species_boot` A data.frame containing the relative abundance indices by region
- `ind_plot` A `ggplot2` object containing the relatie abundance plot 
- `preserve_inter_regional_differences` A logical stating if the relative abundance indices can be compared between regions (if TRUE).

We derive hook-competition-adjusted bootstrap indices using the same function `bootstrap_index_fun_run()`, but adding the argument `ICR_adjust=TRUE`. We derive an overall index and a subregion index (with subregions defined by the survey boundaries).


```{r bootstrap_compfactor, cache=F, warning=F, message=F}
bootstrap_ind_compfactor <-
bootstrap_index_fun_run(species = 'yelloweye',
                          data = Reduced_data_sp,
                          ICR_adjust=T,
                          survey_boundaries = survey_boundaries,
                          R=n_samp,
                          return=T,
                          ncpus=n_cores,
                          type='perc',
                          plot=T
                          )
```

Finally, we compute Poisson-lognormal model-based indices of relative abundance. These include the CPUE-based (unadjusted), hook-competition-adjusted (ICR-based), and censored likelihood-based indices. 

To compute all three indices we use the function `censored_index_fun`. This function takes many arguments, many of which are optional. The non-optional arguments are:

- `data` a `sf` points object containing the IPHC data
- `species` a character name of the species (linked to the variables `N_it` in the dataframe)
- `survey_boundaries` a `sf` polygons object containing the survey boundary definitions
- `M` the number of independent Monte Carlo samples from the posterior used to compute indices
- `ICR_adjust` A logical determining if the ICR-based scale-factor adjustment should be used
- `cprop` the minimum proportion of baits needed to induce censorship. If cprop > 1, then the data are not censored.
- `use_upper_bound` a logical stating if right-censored or interval-censored is desired
- `upper_bound_quantile` a proportion stating which quantiles of observation to remove censorship from. If greater than 1, do not remove censorship from any observation
- `allyears` logical determining if upper bound quantile is computed uniquely for each year (if False) or over all years (if TRUE)
- `preserve_inter_regional_differences` Logical
- `return` logical stating whether or not to return indices as a data.frame
- `plot` logical

The optional arguments are:

- `nthreads` how many cores do you want to run in parallel when fitting the model. Default 1
- `station_effects` logical stating if IID station-station random effects wanted. Default `TRUE`
- `prior_event` a prior distribution for the event-event IID random effect standard deviation (see HT.prior for default half t_7(0,1) distribution)
- `prior_station` a prior distribution for the station-station IID random effect standard deviation (see HT.prior for default half t_7(0,1) distribution)
- `init_vals` initial values for fitting model. If NULL (the default) let INLA choose them.
- `n_knots` the number of knots used for the temporal 'spline'. Default is 8. More knots means a 'wigglier' temporal effect is allowed.
- `seed` the seed used for Monte Carlo sampling. Note that 0 (the default) means the result is non-reproducible, but the code can be significantly faster.
- `verbose` logical. If TRUE, show INLA diagnostics as the model fits. Can help to diagnose issues. Default `FALSE`
- `n_trajectories` integer specifying how many Monte Carlo sampled relative abundance indices to plot on a spaghetti plot. This can help with interpreting uncertainty. Default value 10.
- `keep` a logical stating if you want to return the INLA model object. It can use up memory quickly, but can be useful for diagnosing convergence issues and inspecting model fit. Default `FALSE`

We now fit all three indices

```{r model-based, cache=F, warning=T, message=T, eval=F}
CPUE_ind <-
censored_index_fun(
  data=Reduced_data_sp, survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=1.1, nthreads=n_cores, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, prior_event=HT.prior(), prior_station=HT.prior(), init_vals = c(0.228, -2.346, 2.407, -0.586), n_knots=8, seed=0, verbose=F, n_trajectories=10,
  preserve_inter_regional_differences = F
)

ICR_ind <-
censored_index_fun(
  data=Reduced_data_sp, survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=T, cprop=1.1, nthreads=n_cores, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, prior_event=HT.prior(), prior_station=HT.prior(), init_vals = NULL, n_knots=8, seed=0, verbose=F, n_trajectories=10,
  preserve_inter_regional_differences = F
)

# The following model does not converge
# censored_ind <-
# censored_index_fun(
#   data=Reduced_data_sp, survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=0.95, nthreads=n_cores, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, prior_event=HT.prior(), prior_station=HT.prior(), init_vals = NULL, n_knots=8, seed=0, verbose=F, n_trajectories=10
#     )

# Use upper bound
# Model now converges
censored_ind <-
censored_index_fun(
  data=Reduced_data_sp, survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=0.95, nthreads=n_cores, keep=F, use_upper_bound=TRUE, upper_bound_quantile=0.95, plot=T, allyears=F, station_effects=T, prior_event=HT.prior(), prior_station=HT.prior(), init_vals = c(0.228, -2.346, 2.407, -0.586), n_knots=8, seed=0, verbose=F, n_trajectories=10,
  preserve_inter_regional_differences = F
    )

```

```{r loadprecompilednonspatial, echo=F, message=F, warning=F}
# We don't run the models in the vignette as INLA crashes GitHub Actions
# Instead, we load pre-made files here
list2env(hookCompetition::vignettes_precompiled_data, globalenv())

ggplot2::ggplot(CPUE_inddf, ggplot2::aes(x=.data$year ,y=.data$mean, ymin=.data$q0.025, ymax=.data$q0.975)) +
        ggplot2::geom_point() + ggplot2::geom_errorbar() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~.data$region) +
        ggplot2::ggtitle(paste0('Overdispersed ','Poisson Index ','yelloweye'))

ggplot2::ggplot(ICR_inddf, ggplot2::aes(x=.data$year ,y=.data$mean, ymin=.data$q0.025, ymax=.data$q0.975)) +
        ggplot2::geom_point() + ggplot2::geom_errorbar() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~.data$region) +
        ggplot2::ggtitle(paste0('Overdispersed ','ICR-Adjusted' ,' Poisson Index ','yelloweye'))

ggplot2::ggplot(censored_inddf, ggplot2::aes(x=.data$year ,y=.data$mean, ymin=.data$q0.025, ymax=.data$q0.975)) +
        ggplot2::geom_point() + ggplot2::geom_errorbar() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~.data$region) +
        ggplot2::ggtitle(paste0('Overdispersed ','Censored' ,' Poisson Index ','yelloweye'), subtitle = paste0('censorship proportion ',0.95, ', censorship from data in upper ',max(0,(1-0.95)*100), '% of values removed'))

ggplot2::ggplot(CPUE_trajectory, ggplot2::aes(x=.data$year ,y=.data$mean, colour=.data$MC_ind, group=.data$MC_ind)) +
          ggplot2::geom_line() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~.data$region) +
          ggplot2::ggtitle(paste0('Overdispersed ',' Poisson Index ','yelloweye')) +
          viridis::scale_color_viridis()

ggplot2::ggplot(ICR_trajectory, ggplot2::aes(x=.data$year ,y=.data$mean, colour=.data$MC_ind, group=.data$MC_ind)) +
          ggplot2::geom_line() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~.data$region) +
          ggplot2::ggtitle(paste0('Overdispersed ','ICR-Adjusted', ' Poisson Index ','yelloweye')) +
          viridis::scale_color_viridis()

ggplot2::ggplot(censored_trajectory, ggplot2::aes(x=.data$year ,y=.data$mean, colour=.data$MC_ind, group=.data$MC_ind)) +
          ggplot2::geom_line() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~.data$region) +
          ggplot2::ggtitle(paste0('Overdispersed ','Censored' ,' Poisson Index ','yelloweye'),subtitle = paste0('censorship proportion ',0.95, ', censorship from data in upper ',max(0,(1-0.95)*100), '% of values removed')) +
          viridis::scale_color_viridis()

```


The function `censored_index_fun()` returns a list containing up to 8 entries:

- `pred_overdisp` a data.frame object containing the annual relative abundance indices by region
- `trajectory_samples` a data.frame object containing `n_trajectories` sampled trajectories (time series).
- `posterior_mean` a matrix containing posterior samples of $\hat{\mu}_{i,k}(s, t)$ (i.e. posterior samples of the mean of the counts - useful for posterior predictive checking using the `Bayesplot` package)
- `posterior_sample` a matrix containing posterior predictive samples of $\hat{N}_{i,k}(s, t)$ (i.e. posterior sampled counts from the posterior predictive distribution - useful for posterior predictive checking using the `Bayesplot` package)
- `index_plot` contains the `ggplot2` object of the sampled indices
- `trajectory_plot` contains the `ggplot2` object of the sampled trajectories
- `preserve_inter_regional_differences` a logical stating if the indices are comparable across subregions (if `TRUE`)
- `mod` the returned `INLA` model fit.

Note that the censored Poisson indices can also be fit using the sdmTMB package using the function `censored_index_fun_sdmTMB`. The arguments to the function are similar, with the exclusion of priors and `n_knots` due to penalized splines or first-order random walk structures being placed on the temporal effects. If the relative abundance trend is assumed to be smooth - perhaps for modelling a long-living species with small home range, then set the argument `smoothing_spline=TRUE`. If set to `FALSE`, then a random walk temporal structure is assumed. This allows for larger changes to occur each year (e.g. Pacific Hake summer abundance which changes dramatically from year-year due to summer migration).

Note that the lack of priors penalizing the likelihood on the IID random effects means that we must remove 1996 and 1997 from consideration within the censored Poisson model to make it converge.

We now fit all three indices within sdmTMB

```{r model-based-sdmTMB, cache=F, warning=F, message=F, eval=T}
CPUE_ind_sdmTMB <-
censored_index_fun_sdmTMB(
  data=Reduced_data_sp %>% filter(year>=1998), survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=1.1, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, seed=0, verbose=F, n_trajectories=10,
  preserve_inter_regional_differences = F, smoothing_spline = T
)

ICR_ind_sdmTMB <-
censored_index_fun_sdmTMB(
  data=Reduced_data_sp %>% filter(year>=1998), survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=T, cprop=1.1, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, seed=0, verbose=F, n_trajectories=10,
  preserve_inter_regional_differences = F, smoothing_spline = T
)

censored_ind_sdmTMB <-
censored_index_fun_sdmTMB(
  data=Reduced_data_sp %>% filter(year>=1998), survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=0.95, keep=F, use_upper_bound=TRUE, upper_bound_quantile=0.95, plot=T, allyears=F, station_effects=T, seed=0, verbose=F, n_trajectories=10,
  preserve_inter_regional_differences = F, smoothing_spline = T
    )

```


## Bonus

The package also allows for spatio-temporal indices to be fit

Step 5 load the R objects required for fitting spatio-temporal models. These include: a delauney triangulation mesh for computing the GMRF approximation to the SPDE (cite Lindgren) within r-INLA (cite Rue); an INLA spde object with user-specified pc priors specified on the range and standard deviation parameters. Note that the spde object is a barrier model (cite Bakka) which accounts for land as a barrier.

The function `make_spatial_objects()` achieves this. It takes required arguments:

- `data` a `sf` points object containing the IPHC data
- `survey_boundaries` a `sf` polygons object containing the survey boundary definitions
- `prior_range` Choose a value in kilometers such that the species density at two locations separated by distances greater than this will be `approximately' independent (typically a lower bound is chosen).
- `prior_range_prob` What is you prior probability that the spatial range will be smaller than this value (typically a small probability is chosen to reduce the chances of over-fitting - see the Details section of  ?INLA::inla.spde2.pcmatern for details)
- `prior_sigma` Choose a value for the marginal standard deviation of the random field. Typically an upper bound is used. Remember our linear predictor is on the log scale, so think about constraining exp(+- 2*SD) in terms of multiplicative regional differences. 
- `prior_sigma_prob` What is your prior probability that the true marginal standard deviation exceeds this? (see the Details section of  ?INLA::inla.spde2.pcmatern for details)

It takes optional arguments:

- `mesh_coast` a simplified (smoothed) `sf` polygons object over which the triangulation mesh will be created. If NULL (default), a simplified shapefile of BC Coastline will be returned.
- `hires_coast` a high-resolution `sf` polygons object used for plotting. If NULL (default), a default BC coastline object will be returned.
- `nx` and `ny` How many columns of pixels to place over the survey_boundaries for plotting? Default 150, higher means smoother plots, but longer run times.

```{r makespatialobj, cache=F, warning=F, message=F}
spatial_obj <- make_spatial_objects(
  data=Reduced_data_sp,
  survey_boundaries = survey_boundaries,
  mesh_coast = NULL, hires_coast = NULL, 
  prior_range = 20, prior_range_prob = 0.01,
  prior_sigma = 2, prior_sigma_prob = 0.01 
)

# load it into the global envirionment
list2env(spatial_obj, globalenv())
rm(spatial_obj)

ggplot2::ggplot() + inlabru::gg(mesh) + inlabru::gg(as_Spatial(Reduced_data_sp))

```

The function `make_spatial_objects()` returns a list containing X entries:

- spde_mod an `inla.spde2.pcmatern` object used for setting up the GMRF-SPDE model
- mesh an `inla.mesh.2d` object containing the triangulation mesh used for the random fields
- triangle_ind currently `NULL`
- Coast_hires `sf` polygons containing the high resolution coastline used for plotting
- Coast_simp `sf` polygons containing the low resolution coastline used for building mesh (e.g. no islands or other land barriers)
- predict_pixels `sf` points (pixels) object containing a regular grid of locations across `survey_boundaries` used for plotting and estimating the integrated intensity values for computing indices.

# Step 6 Fit the spatio-temporal indices

The following code *would* be run to make spatio-temporal indices of yelloweye abundance. We do not run the code to save time whilst generating this vignette. Instead, we load pre-compiled objects. Note, we removed map predictions from all years except 2000 to save space.

The function `spatiotemp_censored_index_fun()` is used to fit the spatio-temporal indices. It takes required arguments (the new ones are described):

- `data` a `sf` points object containing the IPHC data
- `species` a character name of the species (linked to the variables `N_it` in the dataframe)
- `survey_boundaries` a `sf` polygons object containing the survey boundary definitions
- `M` the number of independent Monte Carlo samples from the posterior used to compute indices
- `ICR_adjust` A logical determining if the ICR-based scale-factor adjustment should be used
- `cprop` the minimum proportion of baits needed to induce censorship. If cprop > 1, then the data are not censored.
- `use_upper_bound` a logical stating if right-censored or interval-censored is desired
- `upper_bound_quantile` a proportion stating which quantiles of observation to remove censorship from. If greater than 1, do not remove censorship from any observation
- `allyears` logical determining if upper bound quantile is computed uniquely for each year (if False) or over all years (if TRUE)
- `preserve_inter_regional_differences` Logical
- `return` logical stating whether or not to return indices as a data.frame
- `plot` logical
- `mesh` An INLA mesh object (e.g. `mesh` created in `make_spatial_objects()`)
- `spde` An INLA spde object (e.g. `spde_mod` created in `make_spatial_objects()`)
- `pixels` A `sf` points (pixels) object used for plotting the maps of relative abundance


The optional arguments are:

- `nthreads` how many cores do you want to run in parallel when fitting the model. Default 1
- `station_effects` logical stating if IID station-station random effects wanted. Default `TRUE`
- `prior_event` a prior distribution for the event-event IID random effect standard deviation (see HT.prior for default half t_7(0,1) distribution)
- `prior_station` a prior distribution for the station-station IID random effect standard deviation (see HT.prior for default half t_7(0,1) distribution)
- `init_vals` initial values for fitting model. If NULL (the default) let INLA choose them.
- `n_knots` the number of knots used for the temporal 'spline'. Default is 8. More knots means a 'wigglier' temporal effect is allowed.
- `seed` the seed used for Monte Carlo sampling. Note that 0 (the default) means the result is non-reproducible, but the code can be significantly faster.
- `verbose` logical. If TRUE, show INLA diagnostics as the model fits. Can help to diagnose issues. Default `FALSE`
- `n_trajectories` integer specifying how many Monte Carlo sampled relative abundance indices to plot on a spaghetti plot. This can help with interpreting uncertainty. Default value 10.
- `keep` a logical stating if you want to return the INLA model object. It can use up memory quickly, but can be useful for diagnosing convergence issues and inspecting model fit. Default `FALSE`
- `covs` Currently not implemented. Please let me know if (spatio-temporal) environmental covariate functionality is wanted in a future version.
- `spatiotemporal` Logical. Do you want to identify temporal changes in relative abundance uniquely across space? Almost always yes (default `TRUE`).

Note that *these models can take a while to run (5-30 minutes)*. We advise you to set verbose=T so that you can assure yourself that something useful is happening behind the scenes :)

```{r spatiotemporal, eval=F, cache=F, warning=F, message=F}
spatio_temporal_CPUE_ind <-
  spatiotemp_censored_index_fun(data=Reduced_data_sp, survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=1.1, nthreads=n_cores, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, prior_event=HT.prior(), prior_station=HT.prior(), init_vals = c(0.385, -2.083, 2.711, -0.861, 4.000, 0.200, 3.400, 0.300, 2.938), n_knots=8, seed=0, mesh=mesh, spde=spde_mod, pixels=predict_pixels, covs=NULL, spatiotemporal=T, verbose = F, n_trajectories = 10,
  preserve_inter_regional_differences = F)


spatio_temporal_ICR_ind <-
  spatiotemp_censored_index_fun(data=Reduced_data_sp, survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=T, cprop=1.1, nthreads=n_cores, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, prior_event=HT.prior(), prior_station=HT.prior(), init_vals = NULL, n_knots=8, seed=0, mesh=mesh, spde=spde_mod, pixels=predict_pixels, covs=NULL, spatiotemporal=T, verbose = F, n_trajectories = 10,
  preserve_inter_regional_differences = F)

spatio_temporal_censored_ind <-
  spatiotemp_censored_index_fun(data=Reduced_data_sp, survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=0.95, nthreads=n_cores, keep=F, use_upper_bound=TRUE, upper_bound_quantile=0.95, plot=T, allyears=F, station_effects=T, prior_event=HT.prior(), prior_station=HT.prior(), init_vals = NULL, n_knots=8, seed=0, mesh=mesh, spde=spde_mod, pixels=predict_pixels, covs=NULL, spatiotemporal=T, verbose = F, n_trajectories = 10,
  preserve_inter_regional_differences = F)

```

The function `spatiotemp_censored_index_fun()` returns a list containing up to 8 entries:

- `pred_spatiotemp` a data.frame object containing the annual relative abundance indices by region
- `trajectory_samples` a data.frame object containing `n_trajectories` sampled trajectories (time series).
- `posterior_mean` a matrix containing posterior samples of $\hat{\mu}_{i,k}(s, t)$ (i.e. posterior samples of the mean of the counts - useful for posterior predictive checking using the `Bayesplot` package)
- `posterior_sample` a matrix containing posterior predictive samples of $\hat{N}_{i,k}(s, t)$ (i.e. posterior sampled counts from the posterior predictive distribution - useful for posterior predictive checking using the `Bayesplot` package)
- `index_plot` contains the `ggplot2` object of the sampled indices
- `trajectory_plot` contains the `ggplot2` object of the sampled trajectories
- `preserve_inter_regional_differences` a logical stating if the indices are comparable across subregions (if `TRUE`)
- `mod` the returned `INLA` model fit.

Now we plot maps of yelloweye density from the three spatio-temporal models in the year 2000 using the function `spatio_temporal_plot_fun()` which takes arguments:

- `pred_df_plot` the `sf` points (pixels) output named `pred_df_plot` from the function `spatiotemp_censored_index_fun()`
- `year` integer specifying the desired year for plotting
- `variable` character specifying which variable to plot (e.g. 'mean', 'sd', etc.,)
- `hires_COAST` `sf` polygons object with the high resolution coastline
- `plot_figure` Logical - Do you want to plot the figure?
- `return_figure` Logical - Do you want to return the figure?

```{r spatiotemporal_plot, eval=F, message=FALSE, warning=FALSE}
# plot the posterior mean yelloweye density on a map in year 2000
spatio_temporal_plot_fun(
  spatio_temporal_CPUE_plotdf, year=2000, variable='mean', 
  hires_COAST=Coast_hires, plot_figure=T, return_figure=F
)

# plot the posterior mean yelloweye density on a map in year 2000
spatio_temporal_plot_fun(
  spatio_temporal_ICR_plotdf, year=2000, variable='mean', 
  hires_COAST=Coast_hires, plot_figure=T, return_figure=F
)

# plot the posterior mean yelloweye density on a map in year 2000
spatio_temporal_plot_fun(
  spatio_temporal_censored_plotdf, year=2000, variable='mean', 
  hires_COAST=Coast_hires, plot_figure=T, return_figure=F
)

```


```{r loadprecompiledspatiotemp, echo=FALSE, message=FALSE, warning=FALSE, cache=F}
# We don't run the models in the vignette as INLA crashes GitHub Actions
# Instead, we load pre-made files here
ggplot2::ggplot(spatio_temporal_CPUE_inddf, ggplot2::aes(x=year ,y=mean, ymin=q0.025, ymax=q0.975)) +
        ggplot2::geom_point() + ggplot2::geom_errorbar() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~region) +
        ggplot2::ggtitle(paste0('Spatiotemporal Overdispersed Poisson Index ','Yelloweye'))

ggplot2::ggplot(spatio_temporal_ICR_inddf, ggplot2::aes(x=year ,y=mean, ymin=q0.025, ymax=q0.975)) +
        ggplot2::geom_point() + ggplot2::geom_errorbar() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~region) +
                ggplot2::ggtitle(paste0('Spatiotemporal Overdispersed ICR-based Poisson Index ','Yelloweye'))

ggplot2::ggplot(spatio_temporal_censored_inddf, ggplot2::aes(x=year ,y=mean, ymin=q0.025, ymax=q0.975)) +
        ggplot2::geom_point() + ggplot2::geom_errorbar() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~region) +
        ggplot2::ggtitle(paste0('Spatiotemporal Overdispersed Censored Poisson Index yelloweye'), subtitle = 'censorship proportion 0.9, censorship from data in upper 10% of values removed')

ggplot2::ggplot(spatio_temporal_CPUE_trajectory, ggplot2::aes(x=year ,y=mean, colour=MC_ind, group=MC_ind)) +
          ggplot2::geom_line() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~region) +
          ggplot2::ggtitle(paste0('Overdispersed',' Poisson Index ','yelloweye')) +
          viridis::scale_color_viridis()

ggplot2::ggplot(spatio_temporal_ICR_trajectory, ggplot2::aes(x=year ,y=mean, colour=MC_ind, group=MC_ind)) +
          ggplot2::geom_line() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~region) +
          ggplot2::ggtitle(paste0('Overdispersed ICR-based',' Poisson Index ','yelloweye')) +
          viridis::scale_color_viridis()

ggplot2::ggplot(spatio_temporal_censored_trajectory, ggplot2::aes(x=year ,y=mean, colour=MC_ind, group=MC_ind)) +
          ggplot2::geom_line() + ggplot2::ylab('Catch rate index') + ggplot2::facet_grid(~region) +
          ggplot2::ggtitle(paste0('Overdispersed censored',' Poisson Index ','yelloweye'),
                  subtitle = 'Censorship defined as 95% or more baits removed \nCensorship removed from upper 10% of values each year') +          viridis::scale_color_viridis()


```


```{r plotprecompiledspatiotemp, echo=FALSE, message=FALSE, warning=FALSE, cache=FALSE}
# Plot the pre-made files
ggplot2::ggplot() + inlabru::gg(spatio_temporal_CPUE_plotdf[which(spatio_temporal_CPUE_plotdf$year==2000), c('mean')]) + inlabru::gg(sf::as_Spatial(Coast_hires)) + colsc(as.matrix(spatio_temporal_CPUE_plotdf[which(spatio_temporal_CPUE_plotdf$year==2000), c('mean')])) + ggplot2::ggtitle('CPUE-based yelloweye abundance in year 2000')

ggplot2::ggplot() + inlabru::gg(spatio_temporal_ICR_plotdf[which(spatio_temporal_ICR_plotdf$year==2000), c('mean')]) + inlabru::gg(sf::as_Spatial(Coast_hires)) + colsc(as.matrix(spatio_temporal_ICR_plotdf[which(spatio_temporal_ICR_plotdf$year==2000), c('mean')])) + ggplot2::ggtitle('ICR-based yelloweye abundance in year 2000')

ggplot2::ggplot() + inlabru::gg(spatio_temporal_censored_plotdf[which(spatio_temporal_censored_plotdf$year==2000), c('mean')]) + inlabru::gg(sf::as_Spatial(Coast_hires)) + colsc(as.matrix(spatio_temporal_censored_plotdf[which(spatio_temporal_censored_plotdf$year==2000), c('mean')])) + ggplot2::ggtitle('Censored Poisson yelloweye abundance in year 2000',subtitle = 'Censorship defined as 95% or more baits removed \nCensorship removed from upper 10% of values each year')
```

Note that we can also fit spatiotemporal models in sdmTMB using `spatio_temporal_censored_ind_sdmTMB()`. It takes the same arguments with the exception of priors. 

```{r spatiotemporal-sdmTMB, eval=F, cache=F, warning=F, message=F}
spatio_temporal_CPUE_ind_sdmTMB <-
  spatiotemp_censored_index_fun_sdmTMB(data=Reduced_data_sp %>% filter(year>=1998), survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=1.1, keep=T, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, seed=0, mesh=mesh, pixels=predict_pixels, covs=NULL, spatiotemporal='AR1', verbose = T, n_trajectories = 10, preserve_inter_regional_differences = F, grf_priors = sdmTMBpriors(
  matern_s = pc_matern(range_gt = 20, sigma_lt = 2),
  matern_st = pc_matern(range_gt = 20, sigma_lt = 2)),
  smoothing_spline = F)


spatio_temporal_ICR_ind_sdmTMB <-
  spatiotemp_censored_index_fun_sdmTMB(data=Reduced_data_sp %>% filter(year>=1998), survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=T, cprop=1.1, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1.1, plot=T, allyears=F, station_effects=T, seed=0, mesh=mesh, pixels=predict_pixels, covs=NULL, spatiotemporal='AR1', verbose = T, n_trajectories = 10,
  preserve_inter_regional_differences = F, grf_priors = sdmTMBpriors(
  matern_s = pc_matern(range_gt = 20, sigma_lt = 2),
  matern_st = pc_matern(range_gt = 20, sigma_lt = 2)),
  prev_fit = spatio_temporal_CPUE_ind_sdmTMB$mod,
  smoothing_spline = F)

spatio_temporal_censored_ind_sdmTMB <-
  spatiotemp_censored_index_fun_sdmTMB(data=Reduced_data_sp %>% filter(year>=1998), survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=0.95, keep=F, use_upper_bound=TRUE, upper_bound_quantile=0.95, plot=T, allyears=F, station_effects=T, seed=0, mesh=mesh, pixels=predict_pixels, covs=NULL, spatiotemporal='AR1', verbose = T, n_trajectories = 10,
  preserve_inter_regional_differences = F, grf_priors = sdmTMBpriors(
  matern_s = pc_matern(range_gt = 20, sigma_lt = 2),
  matern_st = pc_matern(range_gt = 20, sigma_lt = 2)),
  prev_fit = spatio_temporal_CPUE_ind_sdmTMB$mod,
  smoothing_spline = F)

# Remove the upper bound and upper quantile
spatio_temporal_censored_ind_sdmTMB2 <-
    spatiotemp_censored_index_fun_sdmTMB(data=Reduced_data_sp %>% filter(year>=1998), survey_boundaries=survey_boundaries, species='yelloweye', M=n_samp, return=T, ICR_adjust=F, cprop=0.95, keep=F, use_upper_bound=FALSE, upper_bound_quantile=1, plot=T, allyears=F, station_effects=T, seed=0, mesh=mesh, pixels=predict_pixels, covs=NULL, spatiotemporal='AR1', verbose = T, n_trajectories = 10,
                                         preserve_inter_regional_differences = F, 
                                         grf_priors = sdmTMBpriors(
                                             matern_s = pc_matern(range_gt = 20, sigma_lt = 2),
                                             matern_st = pc_matern(range_gt = 20, sigma_lt = 2)),
                                         prev_fit = spatio_temporal_CPUE_ind_sdmTMB$mod,
  smoothing_spline = F)

```

